---
title: "Penguins ANOVA"
author: "Andrew Paterson"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in pengiuns.csv data
# if you don't save, commit, push, R will override your intermediate versions before they end up on Git

```{r, include = FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, Adelie)?

2. Do some exploring.  This is in tidy format already: tidy format: each column is a different variable, each row is a single observation

New functions: with + table for easy counting over multiple factors

Update species name using case_when

#table() finds counts of things, first separated by Species, second by sex
#with assesses factor levels of a dataframe

```{r}

counts <- with(penguins, table(species, sex))

counts

penguins2 <- penguins %>% 
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie" 
           )
         ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

penguins2

```

-Exploratory Graphs
-Summary statistics
-Levene's tests for equal variances
-