---
title: "Penguins ANOVA"
author: "Andrew Paterson"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in pengiuns.csv data
# if you don't save, commit, push, R will override your intermediate versions before they end up on Git

```{r, include = FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, Adelie)?

2. Do some exploring.  This is in tidy format already: tidy format: each column is a different variable, each row is a single observation

New functions: with + table for easy counting over multiple factors

Update species name using case_when

#table() finds counts of things, first separated by Species, second by sex
#with assesses factor levels of a dataframe

```{r}

counts <- with(penguins, table(species, sex))

counts

penguins2 <- penguins %>% 
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie" 
           )
         ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

penguins2

```

-Exploratory Graphs
-Summary statistics
-Levene's tests for equal variances
-one-way ANOVA
-post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) + 
  facet_wrap(~common_name)
hists

qqplots <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq()+
  facet_wrap(~common_name)

qqplots

#Levene's test for equal variances (> 2 groups... if you have 2 groups, use the f-test for equal variances)

#We'll use the function leveneTest from the 'car' package

#Ho variances are equal
#Ha variances are unequal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# we retain the Ho of equal variances (p = 0.06)

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    Variance = var(flipper_length)
  )

variances

#only at this point is it responsible to use an ANOVA to compare means
```

#3. One-way ANOVA

-single factor is categorgical, has grups within it (species), and one variable we are testing (flipper length)

Single Factor: common_name (species)
Number of levels in the tactor: 3 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for Female Chinstrap, Gentoo, Adelie penguins?

```{r}
#Ho: Mean flipper lengths across all species are equal
#Ha: At least two means differ significantly

#this function says, please compare means for this random variable, grouped by common name

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov) #summary here gives us more info from this outcome


#If we take samples from these three groups and they actually come from ppopulations with the same flipper length, there is almost a zero percent chance we'd get means this or more different by random chance.

#at least two samokes were taken from populations with different means

#There is a significant difference in flipper length based on species. Or, means are not equal across the three different species. Which ones are different?  Are all three different from each other, or something else?

#post-hoc testing using Tukey's HSD
penguins_ph <- TukeyHSD(penguin_aov)
penguins_ph

#Ho is still that means are equal
```

Mean flipper lengths for Adelie, Gentoo, and Chinstrap peguins are all significantly different (F(df) = statistic, p < 0.001, $\alpha$ = .05 with post-hoc Tukey's HSD, $\alpha = 0.05$.

